Working Principles of Proof Assistants and Formalization of some proofs in Agda

### 2025-08-05 Mainpaper/paper.tex
REPORT DONE1

\addcontentsline{toc}{section}{\textbf{ABSTRACT}}  
\clearpage

@@

@@

\tableofcontents
\newpage

@@

\subsection{Objectives}
\begin{enumerate}
    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
    \item To investigate a current existing proof assistants (Agda) and identify the mathematics behind it.
    \item To formalize some selected proofs in Agda and demonstrate verification process.
\end{enumerate}

@@

 The main focus is that this new set of rules does not contain the double negation rule which is present in what we introduced in Section 2.1.1. \textbf {Example \ref{lem}} uses the Double negation rule in its proof, which we don't have now. This agrees to that Law of Excluded Middle does not work in Intuitionstic Logic. The method of proof by contradiction also relies on this rule, so constructivists omit it.\\
These rules are revised with localized hypotheses i.e we use the above rules under a set of premises, and refer the derivation as "derived under a context". With introduction of contexts 
$$Contexts  \ \Gamma ::= .|\Gamma,u: A$$
[See 2.3 \cite{Pfenning2004}]

@@

$\lambda x.t$ is same as $ \lambda y.t[x:=y]$
    \item $\beta$-Reduction\\
     This is substitution. During application all the instances of bound variable in the expression is replaced by the argument. Again referring above example $x^2$ becomes $3^2$.\\
     $(\lambda x.E) F$ is $E[x := F]$, where $E[x := F]$
\end{enumerate}
For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{}

@@

Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.[]
Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values. When inductively defined  datatypes are introduced the
pattern matching becomes even more powerful as patten matching on one value can give the information about another value (thesis)

@@

open import Data.Product using (Sigma; _,_)

@@

\begin{itemize}
    \item In 5.1 we did everything from scratch, starting from definitons by ourselves. For larger and more proofs doing this for every .agda code is difficult. We make use of Agda Standard Library and leverage some pre defined objects and pre proven theorems.
    \item In 5.3, For the second part of the therorem, we had take non constructive approach, which was done by using Dec from Relation.Nullary, which is the law of excluded middle.
    \item In 5.4, We tried to construct real numbers as Cauchy sequences of Rational Numbers like in \cite{BB85}, where we encountered a problem while creating a reciprocal function $\mathbb{N} \to \mathbb{Q}$. The issue being in Agda the numerator requires to be a integer, and we did not find a way to introduce the integer 0. This remains unresolved.
\end{itemize}

% ======== References =========
\bibliographystyle{apalike} 
\bibliography{references.bib}

>>>>

@@

\clearpage
\addcontentsline{toc}{section}{\textbf{ABSTRACT}}

@@

{\small
\tableofcontents
}
\newpage
\section*{Notations}

Throughout this paper, we adopt notation commonly used in type theory and lambda calculus. For clarity, we summarize nonstandard or overloaded symbols below:

\begin{itemize}

  \item $\Gamma \vdash A$ denotes that the judgment $A$ is derivable from the context $\Gamma$. 
  \item We occasionally write $A \subseteq B$ to mean $A$ implies $B$, in the sense of logical entailment. This should not be confused with the subset relation.
  \item $A + B$ denotes the coproduct (disjoint union/ sum) of types $A$ and $B$, given A and B are types.
  \item $\Sigma x : A.\ B(x)$ is the dependent sum type (also written $\sum$), generalizing pairs and existential quantification.
  \item $\Pi x : A.\ B(x)$ denotes the dependent product type (also written $\prod$), generalizing function types and universal quantification.
  \item $\lambda x.\ e$ denotes a function abstraction mapping $x$ to $e$.
  \item $x \mapsto y$ is used informally to denote functions sending $x$ to $y$.
  \item Inference rules and derivations are written using proof trees, with premises above a horizontal bar and conclusions below.
\end{itemize}
\newpage

@@

\subsection{Objectives}
\begin{enumerate}
    \item To explore the correspondence between Proofs and Programs, Type Theory and Intuitionstic Logic.
    \item To investigate current existing proof assistants and identify the mathematics behind it.
    \item To formalize some selected proofs in Agda and demonstrate verification process.
\end{enumerate}

@@

 The main focus is that this new set of rules does not contain the double negation rule which is present in what we introduced in Section 2.1.1. \textbf {Proposition \ref{lem}} uses the Double negation rule in its proof, which we don't have now. This agrees to that Law of Excluded Middle does not work in Intuitionstic Logic. The method of proof by contradiction also relies on this rule, so constructivists omit it.\\
These rules are revised with localized hypotheses i.e we use the above rules under a set of premises, and refer the derivation as "derived under a context". With introduction of contexts 
$$Contexts  \ \Gamma ::= .|\Gamma,u: A$$
[See 2.3 \cite{Pfenning2004}]

@@

$\lambda x.t$ is same as $ \lambda y.t[x:=y]$
    \item $\beta$-Reduction\\
     This is substitution. During application all the instances of bound variable in the expression is replaced by the argument. Again referring above example $x^2$ becomes $3^2$.\\
     $(\lambda x.E) F$ is $E[x := F]$, where $E[x := F]$
\end{enumerate}
For comprehensive treatment of this topic, see \cite{Rojas2015}. And for the original formulation by Church \cite{church1936unsolvable}.

@@

Every agda file has a top level module which corresponds to name of the file and the main programs goes inside the top level module.
Another key component of agda is pattern matching. Pattern matching is a mechanism for analysing the structure of values.

@@

open import Data.Product using (Σ; _,_)

@@

\begin{itemize}
    \item In 5.1 we did everything from scratch, starting from definitons by ourselves. For larger and more proofs doing this for every .agda code is difficult. We make use of Agda Standard Library and leverage some pre defined objects and pre proven theorems.
    \item In 5.3, For the second part of the therorem, we had take non constructive approach, which was done by using Dec from Relation.Nullary, which is the law of excluded middle.
    \item In 5.4, We tried to construct real numbers as Cauchy sequences of Rational Numbers like in \cite{BB85}, where we encountered a problem while creating a reciprocal function $\mathbb{N} \to \mathbb{Q}$. The issue being in Agda the numerator requires to be a integer, and we did not find a way to introduce the integer 0. This remains unresolved.
\end{itemize}
\newpage
% ======== References =========
\bibliographystyle{apalike} 
\bibliography{references.bib}
